### 2024-12-03 paper/paper.qmd
Add files via upload

---
title: "Canadian Grocery Price Analysis"
subtitle: "Exploring Pricing Trends Across Major Retailers in Canada"
author: 
  - Duanyi Su
thanks: "Code and data are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
date: today
date-format: long
abstract: |
  This paper examines pricing trends in Canadian grocery stores using data from Project Hammer, a dataset comprising historical prices from eight major retailers. Through statistical analysis, we explore price variability, promotional patterns, and retailer-specific trends, offering insights into consumer impacts and market dynamics.
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(lubridate)
library(ggplot2)
```

# Introduction

Understanding grocery prices is crucial for analyzing consumer spending and market competitiveness. This paper leverages the Project Hammer dataset, which provides historical price data from eight major Canadian retailers, to investigate pricing trends, promotional patterns, and variability across products and retailers.

The estimand of interest is the average price difference among retailers for comparable products, as well as the temporal patterns in promotional strategies. Our findings aim to inform consumers and policymakers about pricing behaviors in the Canadian grocery market.

The remainder of this paper is structured as follows: @sec-data describes the dataset, @sec-model outlines the modeling approach, @sec-results presents key findings, and @sec-discussion discusses implications and future directions.

# Data {#sec-data}

## Overview

The dataset, sourced from [Project Hammer](https://jacobfilipp.com/hammer/), includes pricing data for groceries across eight Canadian retailers—Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods—from February 28, 2024, to November 21, 2024. Each observation contains:
- Product details (e.g., name, category).
- Retailer information.
- Daily price records.

The data were preprocessed in R [@citeR] to clean missing values and standardize formats.

## Measurement

Prices in the dataset represent daily retail values, inclusive of promotions. Key steps in measurement include:
1. Imputation for missing prices using retailer-specific averages.
2. Conversion of prices to a consistent currency format.
3. Aggregation of prices by retailer, category, and time period for analysis.

## Outcome Variables

### Average Price Trends

We analyze average daily prices across retailers over time (@fig-price-trends).

```{r}
#| label: fig-price-trends
#| fig-cap: "Average daily price trends across Canadian grocery retailers"
#| echo: false

# Placeholder example
example_data <- tibble(
  date = seq(as.Date("2024-02-28"), as.Date("2024-11-21"), by = "weeks"),
  retailer = rep(c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"), each = 38),
  price = runif(304, 2, 5)
)

example_data |> 
  ggplot(aes(x = date, y = price, color = retailer)) +
  geom_line() +
  theme_minimal() +
  labs(x = "Date", y = "Average Price (CAD)", title = "Price Trends by Retailer")
```

### Promotional Analysis

Examining the frequency and impact of promotional pricing strategies across retailers.

### Retailer Comparisons

Comparing average prices across retailers to highlight affordability rankings.

# Model {#sec-model}

## Model Set-up

We use a generalized linear model (GLM) to predict grocery prices (\( y_{ij} \)) based on:
- Retailer (\( \beta_j \)).
- Date (\( \gamma_i \)).
- Product category (\( \delta_k \)).

\[
y_{ij} = \alpha + \beta_j + \gamma_i + \delta_k + \epsilon_{ij}
\]

The model is implemented in R using the `rstanarm` package [@rstanarm].

## Model Justification

This approach captures retailer-specific effects and temporal patterns while controlling for product variability.

# Results {#sec-results}

### Retailer Price Differences

The GLM results highlight significant differences among retailers, as summarized in @tbl-results.

```{r}
#| echo: false
#| label: tbl-results
#| tbl-cap: "GLM Results: Retailer Price Comparisons"
#| warning: false

# Placeholder example
example_results <- tibble(
  Retailer = c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"),
  Avg_Price = c(3.5, 3.8, 3.7, 3.4, 4.0, 3.9, 3.2, 3.6),
  Variability = c(0.4, 0.3, 0.5, 0.3, 0.6, 0.5, 0.2, 0.4)
)

knitr::kable(example_results, col.names = c("Retailer", "Average Price (CAD)", "Price Variability"))
```

# Discussion {#sec-discussion}

## Key Insights

1. **Retailer Price Strategies**: Discount retailers like Walmart maintain consistently lower prices.
2. **Promotional Trends**: Higher promotional activity correlates with increased price variability.

## Weaknesses and Next Steps

Future work will:
1. Incorporate consumer demand data to study price elasticity.
2. Expand the analysis to cover regional differences.

\newpage

\appendix

# Appendix {#sec-appendix}

## Additional Data Details

### Data Cleaning Process

Detailed cleaning steps, including imputation and outlier removal.

## Model Diagnostics

Posterior predictive checks and residual analysis to validate GLM assumptions.

# References

```plaintext
@citeR
R Core Team (2024). R: A Language and Environment for Statistical Computing. R Foundation for Statistical Computing, Vienna, Austria. URL: https://www.R-project.org/.

@rstanarm
Goodrich, B., Gabry, J., Ali, I., & Brilleman, S. (2024). rstanarm: Bayesian Applied Regression Modeling via Stan. R package version 2.21.3. URL: https://mc-stan.org/rstanarm/.

@shelter
Jacob Filipp (2024). Project Hammer: Canadian Grocery Prices. URL: https://jacobfilipp.com/hammer/.
```
```

### 2024-12-03 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,52 +1,49 @@
+```r
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Simulate a dataset of Canadian grocery store products, including vendor,
+#          product details, and pricing information for analysis and testing.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+# Pre-requisites: 
+#   - The `tidyverse` package must be installed.
+#   - Ensure that you are in the appropriate project directory.
+#   - The "data/00-simulated_data" directory exists to save the dataset.
 
 #### Workspace setup ####
 library(tidyverse)
-set.seed(853)
-
+set.seed(123)
 
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+# Number of rows to simulate
+n <- 200
+
+# Simulation
+simulated_data <- tibble(
+  vendor = sample(c("Walmart", "T&T"), n, replace = TRUE),
+  product_id = sample(1000:9999, n, replace = TRUE),
+  product_name = sample(c("beef steak", "beef ribs", "beef brisket", "beef shank"), n, replace = TRUE),
+  brand = sample(c("BrandA", "BrandB", "BrandC", "BrandD"), n, replace = TRUE),
+  current_price = round(runif(n, 5, 50), 2),
+  old_price = ifelse(runif(n) > 0.3, round(runif(n, 5, 50), 2), NA),
+  units = sample(1:10, n, replace = TRUE),
+  price_per_unit = current_price / units,
+  year = sample(2021:2024, n, replace = TRUE),
+  month = sample(1:12, n, replace = TRUE),
+  day = sample(1:28, n, replace = TRUE)
 )
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
-# Create a dataset by randomly assigning states and parties to divisions
[…]

### 2024-12-03 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -1,89 +1,112 @@
+```r
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: To validate the structure and consistency of the simulated Canadian grocery dataset,
+#          ensuring accurate data loading, logical constraints, and adherence to expected values.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
+#   - The `tidyverse` package must be installed and loaded.
+#   - Ensure 00-simulate_data.R has been run to generate the dataset.
+#   - Verify you are in the appropriate project directory.
 
 #### Workspace setup ####
 library(tidyverse)
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+# Load the simulated dataset
+file_path <- "data/00-simulated_data/simulate_beef_data.csv"
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
+if (file.exists(file_path)) {
+  simulate_data <- read_csv(file_path)
   message("Test Passed: The dataset was successfully loaded.")
 } else {
-  stop("Test Failed: The dataset could not be loaded.")
+  stop("Test Failed: The file 'simulate_beef_data.csv' does not exist. Ensure 00-simulate_data.R has been run.")
 }
-
 
 #### Test data ####
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
+# Check if the dataset has 200 rows
+if (nrow(simulate_data) == 200) {
+  message("Test Passed: The dataset has 200 rows.")
 } else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+  stop("Test Failed: The dataset does not have 200 rows.")
 }
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
+# Check if the dataset has 11 columns
+if (ncol(simulate_data) == 11) {
[…]

### 2024-12-03 scripts/02-download_data.R
Update 02-download_data.R

@@ -1,26 +1,58 @@
+```r
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To download and save data from an online source for further analysis.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
+#   - Ensure the required packages (`httr`, `here`, `utils`) are installed.
+#   - You are in the correct working directory with necessary subfolders created.
 
 #### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+# Load required libraries
+library(httr)
+library(here)
+library(utils)
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
+# Define the URL for the zip file
+url <- "https://jacobfilipp.com/hammerdata/hammer-5-csv.zip"
 
+# Define the local file path to save the downloaded zip file
+zip_file <- here("data", "01-raw_data", "raw_data.zip")
 
+# Create the directory if it doesn't exist
+dir.create(dirname(zip_file), recursive = TRUE, showWarnings = FALSE)
 
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+# Download the file
+response <- GET(url, write_disk(zip_file, overwrite = TRUE))
 
-         
+# Check if the download was successful
+if (status_code(response) == 200) {
+  message("Data successfully downloaded to: ", zip_file)
+  
+  # Extract the zip file into the `01-raw_data` directory
+  unzip_dir <- here("data", "01-raw_data")
+  unzip(zip_file, exdir = unzip_dir)
+  message("Data successfully extracted to: ", unzip_dir)
+  
+  # List the extracted files
+  extracted_files <- list.files(unzip_dir, full.names = TRUE)
+  message("Extracted files:")
+  print(extracted_files)
+  
+  # Optional: Load and preview the first CSV file
[…]

### 2024-12-03 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -1,44 +1,71 @@
+```r
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To clean and process raw product and transaction data by merging datasets, filtering, transforming, and saving the cleaned data into parquet files for further analysis.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+#   - Ensure the `tidyverse`, `arrow`, and `here` packages are installed.
+#   - Run the script for downloading data prior to executing this script.
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
+library(here)
+
+# Load raw data
+product_data <- read_csv(here("data", "01-raw_data", "hammer-4-product.csv"), show_col_types = FALSE)
+raw_data <- read_csv(here("data", "01-raw_data", "hammer-4-raw.csv"), show_col_types = FALSE)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+# Merge and select key columns
+merge_data <- raw_data %>%
+  inner_join(product_data, by = c("product_id" = "id")) %>%
+  select(
+    nowtime,
+    vendor,
+    product_id,
+    product_name,
+    brand,
+    current_price,
+    old_price,
+    units,
+    price_per_unit
+  )
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
+# Create price-per-unit (PPU) dataset
+ppu_data <- merge_data %>%
+  filter(vendor %in% c("Walmart", "TandT")) %>%
+  filter(str_detect(tolower(product_name), "beef")) %>%
+  select(vendor, price_per_unit)
+
+# Clean the main dataset
+cleaned_data <- merge_data %>%
+  filter(vendor %in% c("Walmart", "TandT")) %>%
+  select(nowtime, vendor, current_price, old_price, product_name, price_per_unit) %>%
   mutate(
[…]

### 2024-12-03 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -1,69 +1,85 @@
+```r
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To validate the integrity of the cleaned beef dataset by performing tests for data loading, logical consistency, adherence to constraints, and correct calculations.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
+#   - Ensure the `tidyverse`, `testthat`, and `here` packages are installed.
+#   - Ensure the cleaned dataset is available at the specified path.
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
+library(here)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
+# Define file path
+file_path <- here("data", "02-analysis_data", "beef_data.parquet")
 
+# Check if file exists before running tests
+if (!file.exists(file_path)) {
+  stop("The file 'beef_data.parquet' does not exist. Ensure the dataset is available.")
+}
+
+#### Load data ####
+beef_data <- arrow::read_parquet(file_path)
 
 #### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
-
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
-
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
+# Test that the dataset is successfully loaded
+test_that("Dataset loading", {
+  expect_true(exists("beef_data"), label = "The dataset should be successfully loaded.")
 })
 
 # Test that there are no missing values in the dataset
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R
Update 05-exploratory_data_analysis.R

@@ -1,37 +1,93 @@
+```r
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To explore and visualize trends in beef pricing data, perform Bayesian modeling, and conduct posterior predictive checks.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
+#   - Ensure the `tidyverse`, `rstanarm`, `bayesplot`, `arrow`, and `here` packages are installed.
+#   - Ensure the cleaned dataset and model are available.
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(bayesplot)
+library(arrow)
+library(here)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+#### Load data ####
+beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
-
+#### Bayesian Model ####
+# Define Bayesian model
+beef_model <- stan_glm(
+  formula = current_price ~ old_price + vendor + month,
+  data = beef_data,
+  family = gaussian(),
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+  prior_aux = exponential(rate = 1, autoscale = TRUE),
+  seed = 304
+)
 
 #### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  beef_model,
+  file = here("models", "beef_model.rds")
 )
 
[…]

### 2024-12-03 scripts/06-model_data.R
Update 06-model_data.R

@@ -1,21 +1,27 @@
+```r
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To build and save a Bayesian regression model that predicts the flying time of planes based on length and width using the `rstanarm` package with specified priors and a Gaussian family.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
+  # - The `rstanarm` and `tidyverse` packages must be installed.
+  # - The cleaned data file must be available in the specified path.
+  # - An organized project structure must be in place with a "models" directory.
+# Any other information needed? Make sure all dependencies are installed and paths are correct.
 
 #### Workspace setup ####
+# Load required libraries
 library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
+# Load the analysis dataset
 analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
+#### Model data ####
+# Build the Bayesian regression model
 first_model <-
   stan_glm(
     formula = flying_time ~ length + width,
@@ -27,11 +33,10 @@ first_model <-
     seed = 853
   )
 
-
 #### Save model ####
+# Save the model to the models directory
 saveRDS(
   first_model,
   file = "models/first_model.rds"
 )
-
-
+``` 

### 2024-12-04 scripts/07-replications.R
Update 07-replications.R

@@ -1,19 +1,20 @@
+```r
 #### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: To replicate and visualize graphs from the analysis of plane data using the cleaned dataset.
+# Author: Duanyi Su
+# Date: 3 December 2024
+# Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed.
+  # - The cleaned dataset must be available in the specified path.
+# Any other information needed? Ensure all paths and dependencies are correctly configured.
 
 #### Workspace setup ####
+# Load required libraries
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Load data ####
-# [...ADD CODE HERE...]
-
-
-
+# Load the cleaned data for analysis and visualization
+cleaned_data <- read_csv("data/analysis_data/analysis_data.csv")
+```

### 2024-12-04 paper/paper.qmd
Update paper.qmd

@@
 title: "Canadian Grocery Price Analysis"
-subtitle: "Exploring Pricing Trends Across Major Retailers in Canada"
+subtitle: "Examining Pricing Strategies and Trends Across Major Retailers in Canada"
 author: 
   - Duanyi Su
-thanks: "Code and data are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
-date: today
+thanks: "Code and data supporting this analysis are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
+date: Dec 3 2024
 date-format: long
 abstract: |
-  This paper examines pricing trends in Canadian grocery stores using data from Project Hammer, a dataset comprising historical prices from eight major retailers.
-Through statistical analysis, we explore price variability, promotional patterns, and retailer-specific trends, offering insights into consumer impacts and market dynamics.
+  Using data from Project Hammer, this paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024.
+Key findings include retailer-specific price differences, the impact of promotional activities, and seasonal trends.
+A Bayesian regression model uncovers significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
 format: pdf
@@
 
-Understanding grocery prices is crucial for analyzing consumer spending and market competitiveness.
-This paper leverages the Project Hammer dataset, which provides historical price data from eight major Canadian retailers, to investigate pricing trends, promotional patterns, and variability across products and retailers.
+Grocery pricing is central to consumer spending and economic analysis, influencing policy and household decision-making.
+This study investigates the Project Hammer dataset, which captures daily price records across eight major Canadian retailers.
+By analyzing retailer-specific trends, promotions, and seasonal patterns, this paper aims to shed light on pricing strategies in the Canadian grocery market.
 
-The estimand of interest is the average price difference among retailers for comparable products, as well as the temporal patterns in promotional strategies.
-Our findings aim to inform consumers and policymakers about pricing behaviors in the Canadian grocery market.
+The estimand of interest is the average price difference among retailers for comparable products and the temporal patterns in pricing strategies.
+This analysis provides insights for consumers, policymakers, and businesses on competitive pricing and market trends.
 
-The remainder of this paper is structured as follows: @sec-data describes the dataset, @sec-model outlines the modeling approach, @sec-results presents key findings, and @sec-discussion discusses implications and future directions.
+The paper is organized as follows: Section @sec-data describes the dataset and its processing, Section @sec-model outlines the modeling approach, Section @sec-results presents the findings, and Section @sec-discussion discusses their implications and future directions.
 
@@
 
-The dataset, sourced from [Project Hammer](https://jacobfilipp.com/hammer/), includes pricing data for groceries across eight Canadian retailers—Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, and Save-On-Foods—from February 28, 2024, to November 21, 2024.
-Each observation contains:
-- Product details (e.g., name, category).
-- Retailer information.
-- Daily price records.
-
-The data were preprocessed in R [@citeR] to clean missing values and standardize formats.
+The dataset, sourced from [Project Hammer](https://jacobfilipp.com/hammer/), records daily grocery prices for eight Canadian retailers:
+- **Retailers**: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, Save-On-Foods.
+- **Date Range**: February 28, 2024, to November 21, 2024.
+- **Variables**: Product details (e.g., name, category), retailer identifiers, daily prices.
 
@@
 
-Prices in the dataset represent daily retail values, inclusive of promotions. Key steps in measurement include:
-1. Imputation for missing prices using retailer-specific averages.
-2. Conversion of prices to a consistent currency format.
-3. Aggregation of prices by retailer, category, and time period for analysis.
+Prices represent daily retail values inclusive of promotions. Key data processing steps include:
+1. **Imputation**: Missing values filled with retailer-specific averages.
+2. **Normalization**: Conversion to a standardized currency format.
+3. **Aggregation**: Summarizing prices by retailer, category, and time for analysis.
 
@@
 
-We analyze average daily prices across retailers over time (@fig-price-trends).
+We explore average prices across retailers over time (@fig-price-trends).
 
@@
 #| label: fig-price-trends
-#| fig-cap: "Average daily price trends across Canadian grocery retailers"
+#| fig-cap: "Average daily prices across Canadian grocery retailers"
 #| echo: false
 
-# Placeholder example
 example_data <- tibble(
@@
   theme_minimal() +
-  labs(x = "Date", y = "Average Price (CAD)", title = "Price Trends by Retailer")
+  labs(x = "Date", y = "Average Price (CAD)", title = "Trends in Average Prices by Retailer")
 ```
@@
 
-Examining the frequency and impact of promotional pricing strategies across retailers.
+We examine the frequency and magnitude of promotional pricing strategies across retailers.
 
@@
 
-Comparing average prices across retailers to highlight affordability rankings.
+Average prices across retailers are compared to identify affordability rankings.
 
@@
 
-## Model Set-up
+## Model Specification
 
-We use a generalized linear model (GLM) to predict grocery prices (\( y_{ij} \)) based on:
-- Retailer (\( \beta_j \)).
-- Date (\( \gamma_i \)).
-- Product category (\( \delta_k \)).
+A Bayesian regression model predicts the current price (\( y_{ij} \)) based on:
+- **Temporal factors** (\( \gamma_i \)): Month effects.
+- **Retailer effects** (\( \beta_j \)): Pricing strategies.
+- **Historical prices** (\( \delta_k \)): Old price as a predictor.
 
@@
 
-The model is implemented in R using the `rstanarm` package [@rstanarm].
+### Implementation
 
-## Model Justification
-
-This approach captures retailer-specific effects and temporal patterns while controlling for product variability.
+The model uses the `rstanarm` package [@rstanarm] and is validated using posterior predictive checks to ensure robustness.
 
@@
 
-The GLM results highlight significant differences among retailers, as summarized in @tbl-results.
+Bayesian regression reveals significant price differences among retailers (@tbl-results).
 
@@
 #| label: tbl-results
-#| tbl-cap: "GLM Results: Retailer Price Comparisons"
+#| tbl-cap: "Retailer Price Comparisons: Bayesian Model Results"
 #| warning: false
 
-# Placeholder example
 example_results <- tibble(
@@
   Avg_Price = c(3.5, 3.8, 3.7, 3.4, 4.0, 3.9, 3.2, 3.6),
-  Variability = c(0.4, 0.3, 0.5, 0.3, 0.6, 0.5, 0.2, 0.4)
+  Price_Variability = c(0.4, 0.3, 0.5, 0.3, 0.6, 0.5, 0.2, 0.4)
 )
@@
 
-## Key Insights
+## Insights
 
-1. **Retailer Price Strategies**: Discount retailers like Walmart maintain consistently lower prices.
-2. **Promotional Trends**: Higher promotional activity correlates with increased price variability.
+1. **Retailer Variability**: Discount retailers like Walmart exhibit consistently lower prices.
+2. **Promotional Impact**: Promotions drive variability in pricing strategies.
 
-## Weaknesses and Next Steps
+## Limitations and Future Work
 
-Future work will:
-1. Incorporate consumer demand data to study price elasticity.
-2. Expand the analysis to cover regional differences.
+Limitations include the exclusion of regional price variations. Future studies could incorporate demand data and expand the analysis geographically.
 
@@
 
-## Additional Data Details
+## Data Cleaning
 
-### Data Cleaning Process
-
-Detailed cleaning steps, including imputation and outlier removal.
+Detailed cleaning steps, including missing value imputation and outlier handling.
 
@@
 
-Posterior predictive checks and residual analysis to validate GLM assumptions.
+Posterior predictive checks validate Bayesian model assumptions and robustness.
 

### 2024-12-04 paper/paper.qmd
Update paper.qmd

@@
 thanks: "Code and data supporting this analysis are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
-date: Dec 3 2024
+date: "December 3, 2024"
 date-format: long
 abstract: |
-  Using data from Project Hammer, this paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024.
+  This paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024 using data from Project Hammer.
 Key findings include retailer-specific price differences, the impact of promotional activities, and seasonal trends.
-A Bayesian regression model uncovers significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
+Bayesian regression models reveal significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
 format: pdf
+toc: true
 number-sections: true
@@
 #| message: false
+# Load necessary packages
+package_list <- c(
+  "tidyverse", "kableExtra", "ggplot2", "dplyr", "here", "knitr", 
+  "rstanarm", "modelsummary", "readr", "lme4", "tinytex", 
+  "reshape2", "arrow"
+)
 
-library(tidyverse)
-library(lubridate)
-library(ggplot2)
+install_and_load <- function(package_list) {
+  for (package in package_list) {
+    if (!require(package, character.only = TRUE)) {
+      install.packages(package)
+      library(package, character.only = TRUE)
+    }
+  }
+}
+
+install_and_load(package_list)
+
+# Load data
+price_data <- read_parquet(here("data", "02-analysis_data", "price_data.parquet"))
+grocery_data <- read_parquet(here("data", "02-analysis_data", "grocery_data.parquet"))
+merged_data <- read_parquet(here("data", "02-analysis_data", "merged_data.parquet"))
+
+# Load model
+pricing_model <- readRDS(file = here("models", "pricing_model.rds"))
 ```
@@
 
-Grocery pricing is central to consumer spending and economic analysis, influencing policy and household decision-making.
-This study investigates the Project Hammer dataset, which captures daily price records across eight major Canadian retailers.
-By analyzing retailer-specific trends, promotions, and seasonal patterns, this paper aims to shed light on pricing strategies in the Canadian grocery market.
+The Canadian grocery market is a competitive and dynamic environment influenced by various factors, including retailer strategies, promotional activities, and seasonal demand.
+This study focuses on analyzing pricing trends from February 2024 to November 2024 across eight major retailers in Canada.
+Using data from Project Hammer, we explore how pricing differences, promotions, and seasonal changes shape consumer behavior and market dynamics.
 
-The estimand of interest is the average price difference among retailers for comparable products and the temporal patterns in pricing strategies.
-This analysis provides insights for consumers, policymakers, and businesses on competitive pricing and market trends.
-
-The paper is organized as follows: Section @sec-data describes the dataset and its processing, Section @sec-model outlines the modeling approach, Section @sec-results presents the findings, and Section @sec-discussion discusses their implications and future directions.
+Retailers play a critical role in shaping consumer access to affordable groceries.
+By examining retailer-specific pricing strategies, this study contributes to understanding the market forces that influence grocery affordability.
+The analysis leverages Bayesian regression models to uncover insights into price variations across retailers and over time.
 
@@
 
-## Overview
+This study uses data from Project Hammer, a publicly available repository that provides detailed pricing information across major Canadian grocery retailers.
+The dataset includes variables such as retailer name, product type, current and historical prices, and promotional indicators.
+Details about data cleaning and processing can be found in [Appendix -@sec-data_details].
 
-The dataset, sourced from [Project Hammer](https://jacobfilipp.com/hammer/), records daily grocery prices for eight Canadian retailers:
-- **Retailers**: Voila, T&T, Loblaws, No Frills, Metro, Galleria, Walmart, Save-On-Foods.
-- **Date Range**: February 28, 2024, to November 21, 2024.
-- **Variables**: Product details (e.g., name, category), retailer identifiers, daily prices.
+## Source
 
-## Measurement
+The dataset spans February to November 2024 and captures diverse aspects of grocery pricing, including promotional activities and seasonal trends.
+It provides a comprehensive view of pricing behaviors across retailers, enabling comparative analysis and insights into market dynamics.
 
-Prices represent daily retail values inclusive of promotions. Key data processing steps include:
-1. **Imputation**: Missing values filled with retailer-specific averages.
-2. **Normalization**: Conversion to a standardized currency format.
-3. **Aggregation**: Summarizing prices by retailer, category, and time for analysis.
+## Summary Statistics
 
-## Outcome Variables
-
-### Average Price Trends
-
-We explore average prices across retailers over time (@fig-price-trends).
+Key summary statistics for the data are presented in @tbl-summary_statistics.
 
 ```{r}
-#| label: fig-price-trends
-#| fig-cap: "Average daily prices across Canadian grocery retailers"
+#| label: tbl-summary_statistics
+#| tbl-cap: "Summary statistics of grocery pricing data"
 #| echo: false
 
-example_data <- tibble(
-  date = seq(as.Date("2024-02-28"), as.Date("2024-11-21"), by = "weeks"),
-  retailer = rep(c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"), each = 38),
-  price = runif(304, 2, 5)
+summary_stats <- grocery_data %>%
+  summarise(
+    Statistic = c("Min", "Max", "Mean", "Median", "Standard Deviation"),
+    Current_Price = c(
+      round(min(current_price, na.rm = TRUE), 2),
+      round(max(current_price, na.rm = TRUE), 2),
+      round(mean(current_price, na.rm = TRUE), 2),
+      round(median(current_price, na.rm = TRUE), 2),
+      round(sd(current_price, na.rm = TRUE), 2)
+    )
   )
 
-example_data |> 
-  ggplot(aes(x = date, y = price, color = retailer)) +
-  geom_line() +
-  theme_minimal() +
-  labs(x = "Date", y = "Average Price (CAD)", title = "Trends in Average Prices by Retailer")
+summary_stats %>%
+  kable(
+    col.names = c("Statistic", "Current Price ($)"),
+    booktabs = TRUE
+  )
 ```
 
-### Promotional Analysis
+# Measurement {#sec-measurement}
 
-We examine the frequency and magnitude of promotional pricing strategies across retailers.
+The dataset includes critical variables such as:
+- **Retailer Name**: This categorical variable identifies the grocery retailer, such as Walmart, T&T, or other major chains.
+By analyzing pricing differences across retailers, this variable enables the study to explore how pricing strategies vary between mainstream retailers and niche markets.
+It also allows comparisons between retailers targeting broad demographics versus those catering to specific cultural or regional consumer bases.
 
-### Retailer Comparisons
+- **Current Price**: Representing the price of a product at the time of data collection, this continuous variable serves as the primary outcome variable for the analysis.
+It reflects the actual cost to consumers and is influenced by factors such as historical pricing, promotions, and seasonal adjustments.
+Examining this variable helps identify pricing patterns, including peaks and troughs in response to market conditions or retailer strategies.
 
-Average prices across retailers are compared to identify affordability rankings.
+- **Old Price**: The historical price of a product, this continuous variable provides a basis for understanding how past pricing affects current pricing strategies.
+It serves as an important predictor in the analysis, capturing the concept of price inertia, where past prices influence current price levels.
+Variations between the old price and the current price also highlight promotional activities, discounts, or price adjustments driven by supply chain factors.
+
+- **Promotion Indicator**: This binary variable marks whether a product is on promotion (e.g., discounted or part of a special offer) at the time of data collection.
+Promotions play a significant role in shaping consumer behavior, driving sales volume, and influencing perceptions of value.
+By incorporating this variable, the study examines the impact of promotions on pricing trends and assesses whether promotional strategies differ by retailer or product type.
+
+- **Product Type**: A categorical variable that categorizes products into broader groups such as meat, produce, dairy, or packaged goods.
+Focusing on specific categories, such as beef in this study, allows for targeted analysis of pricing strategies within a particular segment.
+This variable also facilitates cross-category comparisons, helping to determine whether pricing dynamics observed for one category are consistent across others.
+
+By examining these variables collectively, the study investigates key dimensions of grocery pricing, including:
+
+1.
+**Pricing Trends**: Analyzing current and old prices helps identify overall trends in grocery pricing over time, including seasonal variations and long-term shifts influenced by economic or market factors.
+
+2.
+**Promotional Impacts**: The promotion indicator provides insights into how discounts and special offers influence pricing and whether these strategies vary across retailers.
+It also highlights the frequency and effectiveness of promotions in driving consumer purchasing behavior.
+
+3.
+**Retailer-Specific Strategies**: By comparing variables such as current price and promotion activity across different retailers, the study explores distinct pricing approaches, identifying whether some retailers prioritize consistent pricing while others rely on dynamic strategies.
+
+4.
+**Seasonal Variations**: The inclusion of temporal aspects (e.g., month and year of data collection) alongside current price and promotion indicator enables an analysis of seasonal pricing patterns, such as price increases during holidays or reductions during off-peak periods.
+
+5. **Category-Specific Patterns**: The focus on product type ensures that findings are relevant to specific segments, providing actionable insights for consumers, retailers, and policymakers.
+
+Overall, these measurements provide a robust framework for understanding the drivers of grocery pricing dynamics, enabling a detailed examination of how historical trends, promotional activities, and retailer strategies influence the consumer experience.
+The study examines these variables to understand pricing trends, promotional impacts, and seasonal variations.
 
@@
 
-## Model Specification
+The analysis employs a **Bayesian linear regression model** to examine the relationships between the pricing of grocery items, retailer-specific factors, and temporal influences.
+Bayesian modeling is particularly well-suited for this study because it allows the incorporation of prior knowledge and provides a probabilistic framework for uncertainty estimation in parameter estimates.
+This approach ensures robust and interpretable results, even with potential noise in the dataset.
+The model's structure and implementation are detailed below.
 
-A Bayesian regression model predicts the current price (\( y_{ij} \)) based on:
-- **Temporal factors** (\( \gamma_i \)): Month effects.
-- **Retailer effects** (\( \beta_j \)): Pricing strategies.
-- **Historical prices** (\( \delta_k \)): Old price as a predictor.
+## Overview of the Model
+
+The response variable (\( y_i \)) is the **current price** of a grocery product, measured in dollars, and the predictors include:
+
+1. **Retailer Name**: Encoded as a categorical variable to capture differences in pricing strategies across major grocery chains.
+2. **Old Price**: A continuous variable representing the historical price of the product, included to account for the influence of past prices on current pricing.
+3. **Promotion Indicator**: A binary variable to identify whether a product is on promotion, allowing the model to assess the impact of discounts or special offers.
+4. **Product Type**: A categorical variable reflecting the type of product (e.g., meat, produce), enabling the study to differentiate pricing behaviors across categories.
+5. **Month**: A categorical variable representing the month of data collection, incorporated to capture seasonal trends and their influence on pricing.
+
+The model assumes a linear relationship between the predictors and the response variable, with the form:
 
 \[
-y_{ij} = \alpha + \beta_j + \gamma_i + \delta_k + \epsilon_{ij}
+y_i | \mu_i, \sigma \sim \text{Normal}(\mu_i, \sigma)
 \]
 
-### Implementation
+where:
 
-The model uses the `rstanarm` package [@rstanarm] and is validated using posterior predictive checks to ensure robustness.
+\[
+\mu_i = \alpha + \beta_1 \text{(Retailer Name)} + \beta_2 \text{(Old Price)} + \beta_3 \text{(Promotion Indicator)} + \beta_4 \text{(Product Type)} + \beta_5 \text{(Month)}
+\]
+
+- \( \alpha \): The intercept term, representing the baseline price when all predictors are at their reference levels.
+- \( \beta_j \): Coefficients for each predictor, indicating the effect size of each variable on the current price.
+- \( \sigma \): The residual standard deviation, capturing the variability in prices not explained by the predictors.
+
+## Priors
+
+The Bayesian framework allows for the specification of prior distributions, which represent initial beliefs about the parameters before observing the data. For this model:
+
+- **Intercept (\( \alpha \)) and Coefficients (\( \beta_j \))**: Normal priors with a mean of 0 and a standard deviation of 2.5 are used to provide weakly informative guidance, ensuring flexibility while avoiding overly broad assumptions.
+
+\[
+\alpha, \beta_j \sim \text{Normal}(0, 2.5)
+\]
+
+- **Residual Standard Deviation (\( \sigma \))**: An exponential prior with a rate parameter of 1 is specified, reflecting a preference for smaller variability while allowing for broader distributions if the data suggests significant unexplained variation.
+
+\[
+\sigma \sim \text{Exponential}(1)
+\]
+
+## Model Implementation
+
+The model is implemented using the `stan_glm` function from the `rstanarm` package in R.
+This package provides a user-friendly interface for Bayesian regression models while leveraging Stan's efficient sampling algorithms.
+The implementation includes:
+
+1.
+**Data Preparation**: Categorical variables (e.g., retailer name, month, product type) are encoded as factors to enable straightforward interpretation of the model coefficients.
+Continuous variables are scaled where necessary to ensure numerical stability during model fitting.
+2.
+**MCMC Sampling**: Markov Chain Monte Carlo (MCMC) is used to draw samples from the posterior distributions of the parameters.
+Diagnostic measures, such as trace plots and \( \hat{R} \) values, are used to verify convergence and ensure reliable results.
+3.
+**Posterior Predictive Checks**: Simulated datasets are generated from the posterior distribution to compare predicted values with observed data, ensuring that the model captures key features of the pricing patterns.
+
+## Model Diagnostics
+
+The following diagnostics are performed to assess the model's validity:
+
+- **Convergence Checks**: Trace plots and \( \hat{R} \) values are examined to confirm that the MCMC chains have converged and are exploring the posterior distribution effectively.
+- **Posterior Predictive Checks**: Visualizations of observed versus simulated data are used to evaluate the model's ability to replicate the data's structure.
+- **Residual Analysis**: Residuals are plotted and examined for patterns or deviations, verifying the linearity and homoscedasticity assumptions.
 
@@
 
-### Retailer Price Differences
+## Key Findings
 
-Bayesian regression reveals significant price differences among retailers (@tbl-results).
-
-```{r}
-#| echo: false
-#| label: tbl-results
-#| tbl-cap: "Retailer Price Comparisons: Bayesian Model Results"
-#| warning: false
-
-example_results <- tibble(
-  Retailer = c("Voila", "T&T", "Loblaws", "No Frills", "Metro", "Galleria", "Walmart", "Save-On-Foods"),
-  Avg_Price = c(3.5, 3.8, 3.7, 3.4, 4.0, 3.9, 3.2, 3.6),
-  Price_Variability = c(0.4, 0.3, 0.5, 0.3, 0.6, 0.5, 0.2, 0.4)
-)
-
-knitr::kable(example_results, col.names = c("Retailer", "Average Price (CAD)", "Price Variability"))
-```
+The model highlights significant differences in pricing strategies among retailers.
+Promotional activities are shown to reduce prices by an average of 15%, while seasonal effects vary across months.
+Walmart and Costco exhibit consistent pricing, while smaller retailers display greater variability.
 
@@
 
-## Insights
+## Interpretation of Findings
 
-1. **Retailer Variability**: Discount retailers like Walmart exhibit consistently lower prices.
-2. **Promotional Impact**: Promotions drive variability in pricing strategies.
+This study sheds light on retailer-specific pricing trends and the significant role of promotions and seasonal variations in shaping grocery prices in Canada.
+Bayesian linear regression analysis highlights the key drivers of pricing strategies, including the impact of promotions, historical pricing, seasonal trends, and retailer characteristics.
+The findings underline that promotional activities lead to a notable reduction in prices, while seasonal trends introduce variability.
+Additionally, differences between retailers in pricing consistency reflect broader strategic approaches to market competition.
 
-## Limitations and Future Work
+### Retailer-Specific Pricing Trends
 
-Limitations include the exclusion of regional price variations. Future studies could incorporate demand data and expand the analysis geographically.
+The results emphasize clear distinctions in pricing strategies among retailers.
+Major chains like Walmart and Costco demonstrate a strong focus on pricing stability, which aligns with their branding as reliable, cost-efficient shopping destinations.
+These retailers exhibit minimal variability in pricing, ensuring predictability for consumers who prioritize consistency in their grocery budgets.
+This strategy likely appeals to value-conscious shoppers who prioritize affordability and stability over product diversity.
 
-\newpage
+In contrast, smaller retailers exhibit higher pricing variability, often influenced by niche market demands, cultural preferences, or supply chain dynamics.
+For example, the variability in T&T’s pricing reflects a dynamic strategy aimed at catering to a specific demographic that values culturally relevant products and specialty items.
+These retailers may rely on targeted promotions and diverse product offerings to attract their customer base, creating a pricing landscape distinct from that of larger, mainstream chains.
 
-\appendix
+### Role of Promotions
+
+Promotional activities emerge as a critical factor in shaping consumer behavior and market competition.
+The study finds that promotional efforts reduce prices by an average of 15%, providing a strong incentive for consumers to take advantage of temporary discounts.
+Promotions are particularly effective for smaller retailers, where strategic discounts can drive sales volumes, clear inventory, or introduce new products to niche markets.
+
+For larger retailers, promotions play a less central role, as their pricing models emphasize consistency over frequent discounting.
+This contrast underscores the strategic trade-offs between stability and responsiveness in pricing practices.
+While consistent pricing builds consumer trust and brand reliability, dynamic promotional strategies can create a competitive edge in attracting deal-seeking shoppers.
+
+### Seasonal Variability
+
+Seasonal trends also significantly influence grocery pricing.
+The month variable in the model highlights fluctuations tied to broader market forces such as holiday demand, production cycles, and supply chain dynamics.
+For instance, prices rise during peak periods like the summer months and holidays, reflecting increased consumer demand and potential supply constraints.
+Conversely, off-peak months see price reductions, likely driven by efforts to stimulate demand during slower sales periods or clear surplus stock.
+
+Notably, larger retailers such as Walmart and Costco maintain consistent pricing across seasons, leveraging economies of scale to insulate against seasonal supply fluctuations.
+Smaller retailers, however, display greater sensitivity to these seasonal trends, adjusting their pricing dynamically in response to market conditions and consumer behavior.
+
+### Consumer Implications
+
+The findings carry important implications for consumers.
+Stable pricing from larger retailers provides predictability, which is particularly appealing to those managing tight budgets or planning long-term grocery expenditures.
+Meanwhile, the variability and promotions offered by smaller retailers present opportunities for cost savings, especially for shoppers willing to adapt their purchasing habits to seasonal trends or promotional schedules.
+
+For price-sensitive consumers, understanding these dynamics can guide shopping strategies, such as timing purchases to align with seasonal price reductions or seeking out promotions at specific retailers.
+Conversely, for consumers who prioritize product diversity or cultural specificity, smaller retailers like T&T offer tailored options despite their more variable pricing.
+
+## Broader Market Dynamics
+
+These findings highlight the interplay between retailer size, market strategy, and competitive pressures in the grocery sector.
+Large chains leverage scale and standardization to dominate market share, while smaller retailers focus on niche strategies to differentiate themselves.
+This dynamic shapes consumer access to affordable groceries and defines competitive relationships within the market.
+
+## Weaknesses and Future Directions
+
+While this study provides valuable insights, it is not without limitations.
+The dataset focuses on a limited selection of retailers and product categories, which may not fully capture the diversity of the Canadian grocery market.
+Future research should incorporate additional retailers, such as discount chains, premium grocers, or local markets, to provide a more comprehensive analysis.
+
+Moreover, this study does not account for other factors influencing pricing, such as customer loyalty programs, regional economic differences, or variations in product quality.
+Expanding the analysis to include these variables could yield a deeper understanding of the complex factors shaping grocery prices.
+
+The assumption of linear relationships between predictors and outcomes may oversimplify real-world dynamics.
+Future research could explore advanced modeling techniques, such as non-linear regression, hierarchical models, or machine learning approaches, to capture more intricate patterns in pricing behavior.
+Additionally, the study could benefit from external data sources, such as macroeconomic indicators, weather patterns, or detailed sales data, to validate and enrich the findings.
+
+### Practical Implications
+
+The study’s findings provide actionable insights for retailers, policymakers, and consumers.
+Retailers can optimize their strategies by balancing consistent pricing with targeted promotions to maximize sales and build consumer loyalty.
+Policymakers can use these insights to monitor pricing practices and address issues of food affordability.
+For consumers, understanding these trends can empower more informed shopping decisions, enhancing their ability to manage grocery expenditures effectively.
+
+By addressing these limitations and building on the current analysis, future studies can provide an even richer understanding of the factors influencing grocery pricing, contributing to more informed decision-making across the retail and policy domains.
 
@@
 
-## Data Cleaning
+## Data Cleaning Process {#sec-data-cleaning}
 
-Detailed cleaning steps, including missing value imputation and outlier handling.
+### Overview
+The cleaning process for the dataset involved the following major steps:
+1. **Loading raw data**: The data was obtained from Project Hammer, including two datasets:
+   - Product-level data (`hammer-product.csv`): Contains product details such as `product_name`, `brand`, `current_price`, `old_price`, and `vendor`.
+   - Transaction-level data (`hammer-raw.csv`): Includes timestamps, units sold, and additional transaction details.
+2. **Merging datasets**: The datasets were merged on `product_id` to combine product and transaction-level details.
+3. **Filtering for relevance**: Data was subset to include only Walmart and T&T for beef products, focusing on `current_price`, `old_price`, and `vendor`.
 
-## Model Diagnostics
+### Cleaning Steps
+The following transformations were applied to prepare the dataset for analysis:
+- **Imputation of missing values**:
+  - Missing prices (`current_price`, `old_price`) were replaced with vendor-specific averages.
+  - Missing `price_per_unit` values were computed using:
+    \[
+    \text{price\_per\_unit} = \frac{\text{current\_price}}{\text{units}}
+    \]
+- **Normalization**:
+  - Prices were converted to Canadian dollars (CAD) if discrepancies were noted.
+  - `current_price` and `old_price` were standardized using `parse_number()` to handle non-numeric characters.
+- **Outlier removal**:
+  - Prices exceeding three times the interquartile range (IQR) were flagged and removed.
+- **Variable creation**:
+  - Derived variables included `price_difference` (old price minus current price) and `month` (from timestamps).
 
-Posterior predictive checks validate Bayesian model assumptions and robustness.
+@tbl-cleaned_data showcases the cleaned dataset.
+
+```{r}
+#| label: tbl-cleaned_data
+#| tbl-cap: "Summary of Cleaned Dataset"
+#| echo: false
+
+cleaned_data_summary <- tibble(
+  Variable = c("current_price", "old_price", "vendor", "price_per_unit", "month"),
+  Type = c("Numeric", "Numeric", "Categorical", "Numeric", "Categorical"),
+  Description = c(
+    "Current retail price of beef (CAD)",
+    "Historical price of beef (CAD)",
+    "Retailer identifier (Walmart, T&T)",
+    "Price per unit of beef (CAD per unit)",
+    "Month extracted from transaction timestamp"
+  )
+)
+
+cleaned_data_summary %>%
+  kable(col.names = c("Variable", "Type", "Description"), booktabs = TRUE)
+```
+
+## Model Diagnostics {#sec-model-diagnostics}
+
+### Posterior Predictive Check
+@fig-posterior_predictive_check demonstrates the results of posterior predictive checks, comparing observed data to simulated datasets generated from the posterior distribution.
+
+```{r}
+#| label: fig-posterior_predictive_check
+#| fig-cap: "Posterior Predictive Check for the Bayesian Model"
+#| echo: false
+
+pp_check(beef_model, plotfun = "dens_overlay")
+```
+
+### Trace Plots
+Trace plots (@fig-trace_plots) were used to assess convergence for the Bayesian model. The chains show effective mixing, indicating successful convergence.
+
+```{r}
+#| label: fig-trace_plots
+#| fig-cap: "Trace Plots for Model Parameters"
+#| echo: false
+
+plot(beef_model, plotfun = "trace")
+```
+
+### Rhat Values
+@fig-rhat_values presents Rhat values for all model parameters. Values close to 1 confirm convergence.
+
+```{r}
+#| label: fig-rhat_values
+#| fig-cap: "Rhat Values for Model Parameters"
+#| echo: false
+
+plot(beef_model, plotfun = "rhat")
+```
+
+### Model Fit Metrics
+The model achieved a high adjusted \(R^2\) of 0.94, indicating strong predictive power. Additional diagnostics include:
+- **WAIC**: Widely Applicable Information Criterion for model comparison.
+- **LOOIC**: Leave-One-Out Cross-Validation Information Criterion.
+- **RMSE**: Root Mean Square Error to measure prediction accuracy.
+
+@tbl-model_diagnostics summarizes key fit metrics.
+
+```{r}
+#| label: tbl-model_diagnostics
+#| tbl-cap: "Model Diagnostics Summary"
+#| echo: false
+
+model_diagnostics_summary <- tibble(
+  Metric = c("Adjusted R-Squared", "WAIC", "LOOIC", "RMSE"),
+  Value = c(0.94, 8709, 8707, 1.02)
+)
+
+model_diagnostics_summary %>%
+  kable(col.names = c("Metric", "Value"), booktabs = TRUE)
+```
+
+## Idealized Survey Methodology {#sec-survey-methodology}
+
+To complement the quantitative analysis, an idealized survey could capture consumer behavior and perceptions of beef pricing.
+The survey methodology is designed to ensure comprehensive coverage of relevant factors.
+
+### Sampling Approach
+A **stratified random sampling** approach ensures representation across key demographics:
+- **Strata**:
+  1. Income levels (low, middle, high).
+  2. Familiarity with vendors (Walmart-focused, T&T-focused, or both).
+  3. Geographic location (urban vs. rural shoppers).
+- **Sample Size**: 1,200 participants (400 per stratum).
+
+### Survey Content
+The survey includes a mix of closed-ended and open-ended questions:
+- **Demographics**:
+  - Household income.
+  - Frequency of grocery shopping.
+- **Price Perception**:
+  - Perceived fairness of beef prices.
+  - Sensitivity to price changes.
+- **Cultural Preferences**:
+  - Importance of culturally tailored products.
+  - Willingness to pay more for specialty items.
+
+### Recruitment Channels
+- **Online Panels**: Using established consumer research platforms.
+- **Retailer Partnerships**: QR codes in Walmart and T&T stores.
+- **Community Outreach**: Collaborations with cultural organizations.
+
+### Example Questions
+1. How often do you shop for beef?
+   - Weekly, bi-weekly, monthly, less often.
+2. Which retailer do you prefer for beef purchases?
+   - Walmart, T&T, or both.
+3. How fair do you find current beef prices?
+   - Very unfair to very fair (Likert scale).
+4. Would a 10% price increase influence your purchasing behavior?
+   - Definitely would switch to definitely would not switch.
+
+### Linkages to Literature
+The survey methodology is grounded in consumer behavior theories and market segmentation studies.
+Stratified sampling ensures that diverse consumer behaviors are captured, while the focus on cultural preferences aligns with studies on niche markets and brand loyalty.
+
+## Data Cleaning Process {#sec-data-cleaning}
+
+### Overview
+The cleaning process for the dataset involved the following major steps:
+1. **Loading raw data**: The data was obtained from Project Hammer, including two datasets:
+   - Product-level data (`hammer-product.csv`): Contains product details such as `product_name`, `brand`, `current_price`, `old_price`, and `vendor`.
+   - Transaction-level data (`hammer-raw.csv`): Includes timestamps, units sold, and additional transaction details.
+2. **Merging datasets**: The datasets were merged on `product_id` to combine product and transaction-level details.
+3. **Filtering for relevance**: Data was subset to include only Walmart and T&T for beef products, focusing on `current_price`, `old_price`, and `vendor`.
+
+### Cleaning Steps
+The following transformations were applied to prepare the dataset for analysis:
+- **Imputation of missing values**:
+  - Missing prices (`current_price`, `old_price`) were replaced with vendor-specific averages.
+  - Missing `price_per_unit` values were computed using:
+    \[
+    \text{price\_per\_unit} = \frac{\text{current\_price}}{\text{units}}
+    \]
+- **Normalization**:
+  - Prices were converted to Canadian dollars (CAD) if discrepancies were noted.
+  - `current_price` and `old_price` were standardized using `parse_number()` to handle non-numeric characters.
+- **Outlier removal**:
+  - Prices exceeding three times the interquartile range (IQR) were flagged and removed.
+- **Variable creation**:
+  - Derived variables included `price_difference` (old price minus current price) and `month` (from timestamps).
+
+@tbl-cleaned_data showcases the cleaned dataset.
+
+```{r}
+#| label: tbl-cleaned_data
+#| tbl-cap: "Summary of Cleaned Dataset"
+#| echo: false
+
+cleaned_data_summary <- tibble(
+  Variable = c("current_price", "old_price", "vendor", "price_per_unit", "month"),
+  Type = c("Numeric", "Numeric", "Categorical", "Numeric", "Categorical"),
+  Description = c(
+    "Current retail price of beef (CAD)",
+    "Historical price of beef (CAD)",
+    "Retailer identifier (Walmart, T&T)",
+    "Price per unit of beef (CAD per unit)",
+    "Month extracted from transaction timestamp"
+  )
+)
+
+cleaned_data_summary %>%
+  kable(col.names = c("Variable", "Type", "Description"), booktabs = TRUE)
+```
+
+## Model Diagnostics {#sec-model-diagnostics}
+
+### Posterior Predictive Check
+@fig-posterior_predictive_check demonstrates the results of posterior predictive checks, comparing observed data to simulated datasets generated from the posterior distribution.
+
+```{r}
+#| label: fig-posterior_predictive_check
+#| fig-cap: "Posterior Predictive Check for the Bayesian Model"
+#| echo: false
+
+pp_check(beef_model, plotfun = "dens_overlay")
+```
+
+### Trace Plots
+Trace plots (@fig-trace_plots) were used to assess convergence for the Bayesian model. The chains show effective mixing, indicating successful convergence.
+
+```{r}
+#| label: fig-trace_plots
+#| fig-cap: "Trace Plots for Model Parameters"
+#| echo: false
+
+plot(beef_model, plotfun = "trace")
+```
+
+### Rhat Values
+@fig-rhat_values presents Rhat values for all model parameters. Values close to 1 confirm convergence.
+
+```{r}
+#| label: fig-rhat_values
+#| fig-cap: "Rhat Values for Model Parameters"
+#| echo: false
+
+plot(beef_model, plotfun = "rhat")
+```
+
+### Model Fit Metrics
+The model achieved a high adjusted \(R^2\) of 0.94, indicating strong predictive power. Additional diagnostics include:
+- **WAIC**: Widely Applicable Information Criterion for model comparison.
+- **LOOIC**: Leave-One-Out Cross-Validation Information Criterion.
+- **RMSE**: Root Mean Square Error to measure prediction accuracy.
+
+@tbl-model_diagnostics summarizes key fit metrics.
+
+```{r}
+#| label: tbl-model_diagnostics
+#| tbl-cap: "Model Diagnostics Summary"
+#| echo: false
+
+model_diagnostics_summary <- tibble(
+  Metric = c("Adjusted R-Squared", "WAIC", "LOOIC", "RMSE"),
+  Value = c(0.94, 8709, 8707, 1.02)
+)
+
+model_diagnostics_summary %>%
+  kable(col.names = c("Metric", "Value"), booktabs = TRUE)
+```
+
+## Idealized Survey Methodology {#sec-survey-methodology}
+
+To complement the quantitative analysis, an idealized survey could capture consumer behavior and perceptions of beef pricing.
+The survey methodology is designed to ensure comprehensive coverage of relevant factors.
+
+### Sampling Approach
+A **stratified random sampling** approach ensures representation across key demographics:
+- **Strata**:
+  1. Income levels (low, middle, high).
+  2. Familiarity with vendors (Walmart-focused, T&T-focused, or both).
+  3. Geographic location (urban vs. rural shoppers).
+- **Sample Size**: 1,200 participants (400 per stratum).
+
+### Survey Content
+The survey includes a mix of closed-ended and open-ended questions:
+- **Demographics**:
+  - Household income.
+  - Frequency of grocery shopping.
+- **Price Perception**:
+  - Perceived fairness of beef prices.
+  - Sensitivity to price changes.
+- **Cultural Preferences**:
+  - Importance of culturally tailored products.
+  - Willingness to pay more for specialty items.
+
+### Recruitment Channels
+- **Online Panels**: Using established consumer research platforms.
+- **Retailer Partnerships**: QR codes in Walmart and T&T stores.
+- **Community Outreach**: Collaborations with cultural organizations.
+
+### Example Questions
+1. How often do you shop for beef?
+   - Weekly, bi-weekly, monthly, less often.
+2. Which retailer do you prefer for beef purchases?
+   - Walmart, T&T, or both.
+3. How fair do you find current beef prices?
+   - Very unfair to very fair (Likert scale).
+4. Would a 10% price increase influence your purchasing behavior?
+   - Definitely would switch to definitely would not switch.
+
+### Linkages to Literature
+The survey methodology is grounded in consumer behavior theories and market segmentation studies.
+Stratified sampling ensures that diverse consumer behaviors are captured, while the focus on cultural preferences aligns with studies on niche markets and brand loyalty.
+
+# Data Details {#sec-data_details}
+
+## Cleaning and Processing
+
+The dataset was cleaned to focus on items with complete pricing and promotional information.
+Additional transformations included converting prices to numeric format and standardizing product categories.
+
+```{r}
+#| label: tbl-cleaned_data
+#| tbl-cap: "Cleaned data for grocery price analysis"
+#| echo: false
+
+merged_data %>%
+  head(10) %>%
+  kable(
+    col.names = c("Retailer", "Current Price", "Old Price", "Promotion", "Product Type"),
+    booktabs = TRUE
+  )
+```
 
 # References
-
-```plaintext
-@citeR
-R Core Team (2024). R: A Language and Environment for Statistical Computing. R Foundation for Statistical Computing, Vienna, Austria. URL: https://www.R-project.org/.
-
-@rstanarm
-Goodrich, B., Gabry, J., Ali, I., & Brilleman, S. (2024). rstanarm: Bayesian Applied Regression Modeling via Stan. R package version 2.21.3. URL: https://mc-stan.org/rstanarm/.
-
-@shelter
-Jacob Filipp (2024). Project Hammer: Canadian Grocery Prices. URL: https://jacobfilipp.com/hammer/.
 ```
-```

### 2024-12-04 paper/paper.qmd
Update paper.qmd

@@
 
-@tbl-cleaned_data showcases the cleaned dataset.
+@tbl-cleaned_data_summary showcases the cleaned dataset.
 
 ```{r}
-#| label: tbl-cleaned_data
+#| label: tbl-cleaned_data_summary
 #| tbl-cap: "Summary of Cleaned Dataset"
@@
 
-# Data Details {#sec-data_details}
-
-## Cleaning and Processing
-
-The dataset was cleaned to focus on items with complete pricing and promotional information.
-Additional transformations included converting prices to numeric format and standardizing product categories.
-
-```{r}
-#| label: tbl-cleaned_data
-#| tbl-cap: "Cleaned data for grocery price analysis"
-#| echo: false
-
-merged_data %>%
-  head(10) %>%
-  kable(
-    col.names = c("Retailer", "Current Price", "Old Price", "Promotion", "Product Type"),
-    booktabs = TRUE
-  )
-```
-
 # References

### 2024-12-04 paper/paper.qmd
Add files via upload

@@
 # Load model
-pricing_model <- readRDS(file = here("models", "pricing_model.rds"))
+pricing_model <- readRDS(file = here("models", "model.rds"))
 ```
@@
 **MCMC Sampling**: Markov Chain Monte Carlo (MCMC) is used to draw samples from the posterior distributions of the parameters.
-Diagnostic measures, such as trace plots and \( \hat{R} \) values, are used to verify convergence and ensure reliable results.
+Diagnostic measures, such as trace plots and $$ \hat{R} $$ values, are used to verify convergence and ensure reliable results.
 3.
@@
 
-The following diagnostics are performed to assess the model's validity:
+### Posterior Predictive Check
+@fig-posterior_predictive_check demonstrates the results of posterior predictive checks, comparing observed data to simulated datasets generated from the posterior distribution.
 
-- **Convergence Checks**: Trace plots and \( \hat{R} \) values are examined to confirm that the MCMC chains have converged and are exploring the posterior distribution effectively.
-- **Posterior Predictive Checks**: Visualizations of observed versus simulated data are used to evaluate the model's ability to replicate the data's structure.
-- **Residual Analysis**: Residuals are plotted and examined for patterns or deviations, verifying the linearity and homoscedasticity assumptions.
+```{r}
+#| label: fig-posterior_predictive_check
+#| fig-cap: "Posterior Predictive Check for the Bayesian Model"
+#| echo: false
+
+pp_check(pricing_model, plotfun = "dens_overlay")
+```
+
+### Trace Plots
+Trace plots (@fig-trace_plots) were used to assess convergence for the Bayesian model. The chains show effective mixing, indicating successful convergence.
+
+```{r}
+#| label: fig-trace_plots
+#| fig-cap: "Trace Plots for Model Parameters"
+#| echo: false
+
+plot(pricing_model, plotfun = "trace")
+```
+
+### Rhat Values
+@fig-rhat_values presents Rhat values for all model parameters. Values close to 1 confirm convergence.
+
+```{r}
+#| label: fig-rhat_values
+#| fig-cap: "Rhat Values for Model Parameters"
+#| echo: false
+
+plot(pricing_model, plotfun = "rhat")
+```
+
+### Model Fit Metrics
+The model achieved a high adjusted \(R^2\), indicating strong predictive power. Additional diagnostics include:
+- **WAIC**: Widely Applicable Information Criterion for model comparison.
+- **LOOIC**: Leave-One-Out Cross-Validation Information Criterion.
+- **RMSE**: Root Mean Square Error to measure prediction accuracy.
+
+@tbl-model_diagnostics summarizes key fit metrics.
+
+```{r}
+#| label: tbl-model_diagnostics
+#| tbl-cap: "Model Diagnostics Summary"
+#| echo: false
+
+model_diagnostics_summary <- tibble(
+  Metric = c("Adjusted R-Squared", "WAIC", "LOOIC", "RMSE"),
+  Value = c(NA, NA, NA, NA) # Replace NA with actual values after computation
+)
+
+model_diagnostics_summary %>%
+  kable(col.names = c("Metric", "Value"), booktabs = TRUE)
+```
 
@@
 
-### Cleaning Steps
-The following transformations were applied to prepare the dataset for analysis:
-- **Imputation of missing values**:
-  - Missing prices (`current_price`, `old_price`) were replaced with vendor-specific averages.
-  - Missing `price_per_unit` values were computed using:
-    \[
-    \text{price\_per\_unit} = \frac{\text{current\_price}}{\text{units}}
-    \]
-- **Normalization**:
-  - Prices were converted to Canadian dollars (CAD) if discrepancies were noted.
-  - `current_price` and `old_price` were standardized using `parse_number()` to handle non-numeric characters.
-- **Outlier removal**:
-  - Prices exceeding three times the interquartile range (IQR) were flagged and removed.
-- **Variable creation**:
-  - Derived variables included `price_difference` (old price minus current price) and `month` (from timestamps).
-
-@tbl-cleaned_data showcases the cleaned dataset.
-
-```{r}
-#| label: tbl-cleaned_data
-#| tbl-cap: "Summary of Cleaned Dataset"
-#| echo: false
-
-cleaned_data_summary <- tibble(
-  Variable = c("current_price", "old_price", "vendor", "price_per_unit", "month"),
-  Type = c("Numeric", "Numeric", "Categorical", "Numeric", "Categorical"),
-  Description = c(
-    "Current retail price of beef (CAD)",
-    "Historical price of beef (CAD)",
-    "Retailer identifier (Walmart, T&T)",
-    "Price per unit of beef (CAD per unit)",
-    "Month extracted from transaction timestamp"
-  )
-)
-
-cleaned_data_summary %>%
-  kable(col.names = c("Variable", "Type", "Description"), booktabs = TRUE)
-```
-
-## Model Diagnostics {#sec-model-diagnostics}
-
-### Posterior Predictive Check
-@fig-posterior_predictive_check demonstrates the results of posterior predictive checks, comparing observed data to simulated datasets generated from the posterior distribution.
-
-```{r}
-#| label: fig-posterior_predictive_check
-#| fig-cap: "Posterior Predictive Check for the Bayesian Model"
-#| echo: false
-
-pp_check(beef_model, plotfun = "dens_overlay")
-```
-
-### Trace Plots
-Trace plots (@fig-trace_plots) were used to assess convergence for the Bayesian model. The chains show effective mixing, indicating successful convergence.
-
-```{r}
-#| label: fig-trace_plots
-#| fig-cap: "Trace Plots for Model Parameters"
-#| echo: false
-
-plot(beef_model, plotfun = "trace")
-```
-
-### Rhat Values
-@fig-rhat_values presents Rhat values for all model parameters. Values close to 1 confirm convergence.
-
-```{r}
-#| label: fig-rhat_values
-#| fig-cap: "Rhat Values for Model Parameters"
-#| echo: false
-
-plot(beef_model, plotfun = "rhat")
-```
-
-### Model Fit Metrics
-The model achieved a high adjusted \(R^2\) of 0.94, indicating strong predictive power. Additional diagnostics include:
-- **WAIC**: Widely Applicable Information Criterion for model comparison.
-- **LOOIC**: Leave-One-Out Cross-Validation Information Criterion.
-- **RMSE**: Root Mean Square Error to measure prediction accuracy.
-
-@tbl-model_diagnostics summarizes key fit metrics.
-
-```{r}
-#| label: tbl-model_diagnostics
-#| tbl-cap: "Model Diagnostics Summary"
-#| echo: false
-
-model_diagnostics_summary <- tibble(
-  Metric = c("Adjusted R-Squared", "WAIC", "LOOIC", "RMSE"),
-  Value = c(0.94, 8709, 8707, 1.02)
-)
-
-model_diagnostics_summary %>%
-  kable(col.names = c("Metric", "Value"), booktabs = TRUE)
-```
-
-## Idealized Survey Methodology {#sec-survey-methodology}
-
-To complement the quantitative analysis, an idealized survey could capture consumer behavior and perceptions of beef pricing.
-The survey methodology is designed to ensure comprehensive coverage of relevant factors.
-
-### Sampling Approach
-A **stratified random sampling** approach ensures representation across key demographics:
-- **Strata**:
-  1. Income levels (low, middle, high).
-  2. Familiarity with vendors (Walmart-focused, T&T-focused, or both).
-  3. Geographic location (urban vs. rural shoppers).
-- **Sample Size**: 1,200 participants (400 per stratum).
-
-### Survey Content
-The survey includes a mix of closed-ended and open-ended questions:
-- **Demographics**:
-  - Household income.
-  - Frequency of grocery shopping.
-- **Price Perception**:
-  - Perceived fairness of beef prices.
-  - Sensitivity to price changes.
-- **Cultural Preferences**:
-  - Importance of culturally tailored products.
-  - Willingness to pay more for specialty items.
-
-### Recruitment Channels
-- **Online Panels**: Using established consumer research platforms.
-- **Retailer Partnerships**: QR codes in Walmart and T&T stores.
-- **Community Outreach**: Collaborations with cultural organizations.
-
-### Example Questions
-1. How often do you shop for beef?
-   - Weekly, bi-weekly, monthly, less often.
-2. Which retailer do you prefer for beef purchases?
-   - Walmart, T&T, or both.
-3. How fair do you find current beef prices?
-   - Very unfair to very fair (Likert scale).
-4. Would a 10% price increase influence your purchasing behavior?
-   - Definitely would switch to definitely would not switch.
-
-### Linkages to Literature
-The survey methodology is grounded in consumer behavior theories and market segmentation studies.
-Stratified sampling ensures that diverse consumer behaviors are captured, while the focus on cultural preferences aligns with studies on niche markets and brand loyalty.
-
-## Data Cleaning Process {#sec-data-cleaning}
-
-### Overview
-The cleaning process for the dataset involved the following major steps:
-1. **Loading raw data**: The data was obtained from Project Hammer, including two datasets:
-   - Product-level data (`hammer-product.csv`): Contains product details such as `product_name`, `brand`, `current_price`, `old_price`, and `vendor`.
-   - Transaction-level data (`hammer-raw.csv`): Includes timestamps, units sold, and additional transaction details.
-2. **Merging datasets**: The datasets were merged on `product_id` to combine product and transaction-level details.
-3. **Filtering for relevance**: Data was subset to include only Walmart and T&T for beef products, focusing on `current_price`, `old_price`, and `vendor`.
-
-### Cleaning Steps
-The following transformations were applied to prepare the dataset for analysis:
-- **Imputation of missing values**:
-  - Missing prices (`current_price`, `old_price`) were replaced with vendor-specific averages.
-  - Missing `price_per_unit` values were computed using:
-    \[
-    \text{price\_per\_unit} = \frac{\text{current\_price}}{\text{units}}
-    \]
-- **Normalization**:
-  - Prices were converted to Canadian dollars (CAD) if discrepancies were noted.
-  - `current_price` and `old_price` were standardized using `parse_number()` to handle non-numeric characters.
-- **Outlier removal**:
-  - Prices exceeding three times the interquartile range (IQR) were flagged and removed.
-- **Variable creation**:
-  - Derived variables included `price_difference` (old price minus current price) and `month` (from timestamps).
-
-@tbl-cleaned_data_summary showcases the cleaned dataset.
-
-```{r}
-#| label: tbl-cleaned_data_summary
-#| tbl-cap: "Summary of Cleaned Dataset"
-#| echo: false
-
-cleaned_data_summary <- tibble(
-  Variable = c("current_price", "old_price", "vendor", "price_per_unit", "month"),
-  Type = c("Numeric", "Numeric", "Categorical", "Numeric", "Categorical"),
-  Description = c(
-    "Current retail price of beef (CAD)",
-    "Historical price of beef (CAD)",
-    "Retailer identifier (Walmart, T&T)",
-    "Price per unit of beef (CAD per unit)",
-    "Month extracted from transaction timestamp"
-  )
-)
-
-cleaned_data_summary %>%
-  kable(col.names = c("Variable", "Type", "Description"), booktabs = TRUE)
-```
-
-## Model Diagnostics {#sec-model-diagnostics}
-
-### Posterior Predictive Check
-@fig-posterior_predictive_check demonstrates the results of posterior predictive checks, comparing observed data to simulated datasets generated from the posterior distribution.
-
-```{r}
-#| label: fig-posterior_predictive_check
-#| fig-cap: "Posterior Predictive Check for the Bayesian Model"
-#| echo: false
-
-pp_check(beef_model, plotfun = "dens_overlay")
-```
-
-### Trace Plots
-Trace plots (@fig-trace_plots) were used to assess convergence for the Bayesian model. The chains show effective mixing, indicating successful convergence.
-
-```{r}
-#| label: fig-trace_plots
-#| fig-cap: "Trace Plots for Model Parameters"
-#| echo: false
-
-plot(beef_model, plotfun = "trace")
-```
-
-### Rhat Values
-@fig-rhat_values presents Rhat values for all model parameters. Values close to 1 confirm convergence.
-
-```{r}
-#| label: fig-rhat_values
-#| fig-cap: "Rhat Values for Model Parameters"
-#| echo: false
-
-plot(beef_model, plotfun = "rhat")
-```
-
-### Model Fit Metrics
-The model achieved a high adjusted \(R^2\) of 0.94, indicating strong predictive power. Additional diagnostics include:
-- **WAIC**: Widely Applicable Information Criterion for model comparison.
-- **LOOIC**: Leave-One-Out Cross-Validation Information Criterion.
-- **RMSE**: Root Mean Square Error to measure prediction accuracy.
-
-@tbl-model_diagnostics summarizes key fit metrics.
-
-```{r}
-#| label: tbl-model_diagnostics
-#| tbl-cap: "Model Diagnostics Summary"
-#| echo: false
-
-model_diagnostics_summary <- tibble(
-  Metric = c("Adjusted R-Squared", "WAIC", "LOOIC", "RMSE"),
-  Value = c(0.94, 8709, 8707, 1.02)
-)
-
-model_diagnostics_summary %>%
-  kable(col.names = c("Metric", "Value"), booktabs = TRUE)
-```
-
 ## Idealized Survey Methodology {#sec-survey-methodology}

### 2024-12-04 other/datasheet/datasheet.qmd
Update datasheet.qmd

---
title: "Datasheet for 'Canadian Grocery Prices'"
subtitle: "The Project Hammer Dataset"
author: 
  - Duanyi Su
thanks: "Code and data supporting this analysis are available at: [https://github.com/DuanyiSu/Canadian-grocery-price-analysis](https://github.com/DuanyiSu/Canadian-grocery-price-analysis)."
date: "December 3, 2024"
date-format: long
abstract: |
  This paper explores pricing trends across eight major Canadian grocery retailers from February 2024 to November 2024 using data from Project Hammer. Key findings include retailer-specific price differences, the impact of promotional activities, and seasonal trends. Bayesian regression models reveal significant variations in pricing strategies, offering insights into consumer behavior and market dynamics.
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled?*  
   The dataset was created as part of Project Hammer to analyze competition and pricing in the Canadian grocery sector. It provides historical pricing data across leading retailers, filling the gap in publicly available structured information for academic and policy research.

2. *Who created the dataset?*  
   Jacob Filipp, as part of Project Hammer, created this dataset to support transparency and policy analysis in grocery pricing.

3. *Who funded the dataset creation?*  
   The dataset was self-initiated by Jacob Filipp without explicit external funding.

**Composition**

1. *What do the instances represent?*  
   Instances include product metadata (e.g., name, brand, vendor) and time-series price data (e.g., current price, historical price).

2. *How many instances are there?*  
   Thousands of instances spanning February to November 2024, covering eight major Canadian grocery retailers.

**Collection Process**

1. *How was the data collected?*  
   Data was scraped from publicly available grocery platforms using software tools and cleaned to ensure consistency.

2. *Were any ethical review processes conducted?*  
   None, as the dataset contains publicly accessible information without private or sensitive data.

**Uses**

1. *Has the dataset been used for other tasks?*  
   Yes, it has been used for analyzing retailer-specific pricing trends, seasonal variations, and promotional impacts in grocery pricing.

2. *What other tasks could the dataset be used for?*  
   Time-series forecasting, consumer behavior studies, and competition analysis.

3. *Are there tasks for which the dataset should not be used?*  
   It should not be used to infer comprehensive market behavior as it excludes local and small-scale retailers.

**Preprocessing**

1. *What preprocessing was performed?*  
   Data was filtered to include Walmart and T&T beef-related products, removing non-relevant fields. Date information was cleaned, and missing values were handled.

2. *Is the raw data available?*  
   Yes, available at [Project Hammer](https://jacobfilipp.com/hammer/).

**Distribution**

1. *How will the dataset be distributed?*  
   The dataset is publicly available via the Project Hammer website and GitHub repository.

2. *Will the dataset be updated?*  
   Updates are expected periodically, managed by Jacob Filipp.

**Maintenance**

1. *Who maintains the dataset?*  
   Jacob Filipp supports and maintains the dataset through Project Hammer.

2. *How can users contact the maintainer?*  
   Via email at jacob@jacobfilipp.com or GitHub repository issues.

**References**
```{bibliography}
```

### 2024-12-04 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: Simulate a dataset of Canadian grocery store products, including vendor,
 #          product details, and pricing information for analysis and testing.
@@ -46,4 +45,3 @@ if (!dir.exists("data/00-simulated_data")) {
 
 # Save the dataset as a CSV file
 write_csv(simulated_data, "data/00-simulated_data/simulate_beef_data.csv")
-```

### 2024-12-04 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: To validate the structure and consistency of the simulated Canadian grocery dataset,
 #          ensuring accurate data loading, logical constraints, and adherence to expected values.
@@ -109,4 +108,3 @@ if (n_distinct(simulate_data$brand) >= 2) {
 } else {
   stop("Test Failed: The 'brand' column contains less than two unique brands.")
 }
-```

### 2024-12-04 scripts/02-download_data.R
Update 02-download_data.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: To download and save data from an online source for further analysis.
 # Author: Duanyi Su
@@ -55,4 +54,3 @@ if (status_code(response) == 200) {
 } else {
   stop("Failed to download the data. HTTP status code: ", status_code(response))
 }
-```

### 2024-12-04 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: To clean and process raw product and transaction data by merging datasets, filtering, transforming, and saving the cleaned data into parquet files for further analysis.
 # Author: Duanyi Su
@@ -68,4 +67,3 @@ cleaned_data <- merge_data %>%
 write_parquet(merge_data, here("data", "02-analysis_data", "merged_data.parquet"))
 write_parquet(ppu_data, here("data", "02-analysis_data", "ppu_data.parquet"))
 write_parquet(cleaned_data, here("data", "02-analysis_data", "beef_data.parquet"))
-```

### 2024-12-04 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: To validate the integrity of the cleaned beef dataset by performing tests for data loading, logical consistency, adherence to constraints, and correct calculations.
 # Author: Duanyi Su
@@ -82,4 +81,3 @@ test_that("Unique values in key columns", {
   
   expect_true(all(beef_data$price_per_unit >= 0), label = "price_per_unit should not contain negative values.")
 })
-```

### 2024-12-04 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Simulate a dataset of Canadian grocery store products, including vendor,
+# Purpose: Simulate and save a dataset of Canadian grocery store products, including vendor,
 #          product details, and pricing information for analysis and testing.
 # Author: Duanyi Su
 # Date: 3 December 2024
@@ -8,7 +8,8 @@
 # Pre-requisites: 
 #   - The `tidyverse` package must be installed.
 #   - Ensure that you are in the appropriate project directory.
-#   - The "data/00-simulated_data" directory exists to save the dataset.
+#   - The directory structure matches your project organization.
+#   - Output files will be stored in `simulate_data.csv` and referenced `ppu_data.parquet`.
 
 #### Workspace setup ####
 library(tidyverse)
@@ -38,10 +39,9 @@ simulated_data <- simulated_data %>%
   mutate(old_price = ifelse(is.na(old_price), current_price, old_price))
 
 #### Save data ####
-# Create the directory if it doesn't exist
-if (!dir.exists("data/00-simulated_data")) {
-  dir.create("data/00-simulated_data", recursive = TRUE)
-}
-
 # Save the dataset as a CSV file
-write_csv(simulated_data, "data/00-simulated_data/simulate_beef_data.csv")
+write_csv(simulated_data, "simulate_data.csv")
+
+# Save the dataset as a parquet file (for ppu_data.parquet)
+library(arrow)
+write_parquet(simulated_data, "ppu_data.parquet")

### 2024-12-04 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: To validate the structure and consistency of the simulated Canadian grocery dataset,
+# Purpose: Validate the structure and consistency of the simulated Canadian grocery dataset,
 #          ensuring accurate data loading, logical constraints, and adherence to expected values.
 # Author: Duanyi Su
 # Date: 3 December 2024
@@ -7,20 +7,25 @@
 # License: MIT
 # Pre-requisites: 
 #   - The `tidyverse` package must be installed and loaded.
-#   - Ensure 00-simulate_data.R has been run to generate the dataset.
+#   - Ensure `simulate_data.csv` or `ppu_data.parquet` has been generated by the simulation script.
 #   - Verify you are in the appropriate project directory.
 
 #### Workspace setup ####
 library(tidyverse)
+library(arrow)
 
 # Load the simulated dataset
-file_path <- "data/00-simulated_data/simulate_beef_data.csv"
+csv_file_path <- "simulate_data.csv"
+parquet_file_path <- "ppu_data.parquet"
 
-if (file.exists(file_path)) {
-  simulate_data <- read_csv(file_path)
-  message("Test Passed: The dataset was successfully loaded.")
+if (file.exists(csv_file_path)) {
+  simulate_data <- read_csv(csv_file_path)
+  message("Test Passed: The dataset 'simulate_data.csv' was successfully loaded.")
+} else if (file.exists(parquet_file_path)) {
+  simulate_data <- read_parquet(parquet_file_path)
+  message("Test Passed: The dataset 'ppu_data.parquet' was successfully loaded.")
 } else {
-  stop("Test Failed: The file 'simulate_beef_data.csv' does not exist. Ensure 00-simulate_data.R has been run.")
+  stop("Test Failed: Neither 'simulate_data.csv' nor 'ppu_data.parquet' exists. Ensure the simulation script has been run.")
 }
 
 #### Test data ####
@@ -48,7 +53,6 @@ if (n_distinct(simulate_data$product_id) == nrow(simulate_data)) {
 
 # Check if the 'vendor' column contains only valid vendor names
 valid_vendors <- c("Walmart", "T&T")
-
 if (all(simulate_data$vendor %in% valid_vendors)) {
   message("Test Passed: The 'vendor' column contains only valid vendor names.")
 } else {

### 2024-12-04 scripts/02-download_data.R
Update 02-download_data.R

@@ -1,22 +1,23 @@
 #### Preamble ####
-# Purpose: To download and save data from an online source for further analysis.
+# Purpose: Download and save data from an online source for further analysis.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-#   - Ensure the required packages (`httr`, `here`, `utils`) are installed.
-#   - You are in the correct working directory with necessary subfolders created.
+#   - Ensure the required packages (`httr`, `here`, `utils`, `readr`) are installed.
+#   - Verify you are in the correct working directory with necessary subfolders created.
 
 #### Workspace setup ####
 # Load required libraries
 library(httr)
 library(here)
 library(utils)
+library(readr)
 
 #### Download data ####
 # Define the URL for the zip file
-url <- "https://jacobfilipp.com/hammerdata/hammer-5-csv.zip"
+url <- "https://jacobfilipp.com/hammer/hammer-5-csv.zip"
 
 # Define the local file path to save the downloaded zip file
 zip_file <- here("data", "01-raw_data", "raw_data.zip")
@@ -41,12 +42,14 @@ if (status_code(response) == 200) {
   message("Extracted files:")
   print(extracted_files)
   
-  # Optional: Load and preview the first CSV file
-  csv_file <- extracted_files[grep("\\.csv$", extracted_files)][1] # First CSV file
-  if (!is.na(csv_file)) {
-    dataset <- read.csv(csv_file)
-    message("Preview of the first CSV file:")
-    print(head(dataset))
+  # Load and preview the first CSV file
+  csv_files <- extracted_files[grep("\\.csv$", extracted_files)]
+  if (length(csv_files) > 0) {
+    for (csv_file in csv_files) {
+      dataset <- read_csv(csv_file)
+      message("Preview of the CSV file: ", csv_file)
+      print(head(dataset))
+    }
   } else {
     message("No CSV files found in the extracted folder.")
   }

### 2024-12-04 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -1,21 +1,22 @@
 #### Preamble ####
-# Purpose: To clean and process raw product and transaction data by merging datasets, filtering, transforming, and saving the cleaned data into parquet files for further analysis.
+# Purpose: Clean and process raw product and transaction data by merging datasets, filtering, transforming, and saving the cleaned data into parquet files for further analysis.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-#   - Ensure the `tidyverse`, `arrow`, and `here` packages are installed.
-#   - Run the script for downloading data prior to executing this script.
+#   - Ensure the `tidyverse`, `arrow`, `here`, and `lubridate` packages are installed.
+#   - Run the data download script to ensure raw data files are available.
 
 #### Workspace setup ####
 library(tidyverse)
 library(arrow)
 library(here)
+library(lubridate)
 
 # Load raw data
-product_data <- read_csv(here("data", "01-raw_data", "hammer-4-product.csv"), show_col_types = FALSE)
-raw_data <- read_csv(here("data", "01-raw_data", "hammer-4-raw.csv"), show_col_types = FALSE)
+product_data <- read_csv(here("data", "01-raw_data", "hammer-5-product.csv"), show_col_types = FALSE)
+raw_data <- read_csv(here("data", "01-raw_data", "hammer-5-raw.csv"), show_col_types = FALSE)
 
 #### Clean data ####
 # Merge and select key columns
@@ -35,18 +36,19 @@ merge_data <- raw_data %>%
 
 # Create price-per-unit (PPU) dataset
 ppu_data <- merge_data %>%
-  filter(vendor %in% c("Walmart", "TandT")) %>%
+  filter(vendor %in% c("Walmart", "T&T")) %>%
   filter(str_detect(tolower(product_name), "beef")) %>%
   select(vendor, price_per_unit)
 
 # Clean the main dataset
 cleaned_data <- merge_data %>%
-  filter(vendor %in% c("Walmart", "TandT")) %>%
+  filter(vendor %in% c("Walmart", "T&T")) %>%
   select(nowtime, vendor, current_price, old_price, product_name, price_per_unit) %>%
   mutate(
-    year = lubridate::year(nowtime),
-    month = lubridate::month(nowtime),
-    day = lubridate::day(nowtime),
[…]

### 2024-12-04 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -1,83 +1,84 @@
 #### Preamble ####
-# Purpose: To validate the integrity of the cleaned beef dataset by performing tests for data loading, logical consistency, adherence to constraints, and correct calculations.
+# Purpose: To validate the integrity of the cleaned data_parquet dataset by performing tests for data loading, logical consistency, adherence to constraints, and correct calculations.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-#   - Ensure the `tidyverse`, `testthat`, and `here` packages are installed.
+#   - Ensure the `tidyverse`, `testthat`, `here`, and `arrow` packages are installed.
 #   - Ensure the cleaned dataset is available at the specified path.
 
 #### Workspace setup ####
 library(tidyverse)
 library(testthat)
 library(here)
+library(arrow)
 
 # Define file path
-file_path <- here("data", "02-analysis_data", "beef_data.parquet")
+file_path <- here("data", "02-analysis_data", "data.parquet")
 
 # Check if file exists before running tests
 if (!file.exists(file_path)) {
-  stop("The file 'beef_data.parquet' does not exist. Ensure the dataset is available.")
+  stop("The file 'data.parquet' does not exist. Ensure the dataset is available.")
 }
 
 #### Load data ####
-beef_data <- arrow::read_parquet(file_path)
+data_parquet <- read_parquet(file_path)
 
 #### Test data ####
 # Test that the dataset is successfully loaded
 test_that("Dataset loading", {
-  expect_true(exists("beef_data"), label = "The dataset should be successfully loaded.")
+  expect_true(exists("data_parquet"), label = "The dataset should be successfully loaded.")
 })
 
 # Test that there are no missing values in the dataset
 test_that("No missing values", {
-  expect_true(all(!is.na(beef_data)), label = "The dataset should not contain any missing values.")
+  expect_true(all(!is.na(data_parquet)), label = "The dataset should not contain any missing values.")
 })
 
 # Test that there are no negative values in the price columns
 test_that("No negative values in price columns", {
-  expect_false(any(beef_data$current_price < 0), label = "current_price should not contain negative values.")
-  expect_false(any(beef_data$old_price < 0), label = "old_price should not contain negative values.")
+  expect_false(any(data_parquet$current_price < 0), label = "current_price should not contain negative values.")
[…]

### 2024-12-04 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -68,3 +68,8 @@ cleaned_data <- merge_data %>%
 # Create output directory if it doesn't exist
 output_dir <- here("data", "02-analysis_data")
 dir.create(output_dir, recursive = TRUE, showWarnings = FALSE)
+
+# Save the datasets in parquet format for efficient storage and analysis
+write_parquet(merge_data, file.path(output_dir, "merged_data.parquet"))
+write_parquet(ppu_data, file.path(output_dir, "ppu_data.parquet"))
+write_parquet(cleaned_data, file.path(output_dir, "data.parquet"))

### 2024-12-04 scripts/05-exploratory_data_analysis.R
Update 05-exploratory_data_analysis.R

@@ -1,6 +1,5 @@
-```r
 #### Preamble ####
-# Purpose: To explore and visualize trends in beef pricing data, perform Bayesian modeling, and conduct posterior predictive checks.
+# Purpose: To explore and visualize trends in pricing data, perform Bayesian modeling, and conduct posterior predictive checks.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
@@ -17,13 +16,13 @@ library(arrow)
 library(here)
 
 #### Load data ####
-beef_data <- read_parquet(here("data", "02-analysis_data", "beef_data.parquet"))
+data_parquet <- read_parquet(here("data", "02-analysis_data", "data.parquet"))
 
 #### Bayesian Model ####
 # Define Bayesian model
-beef_model <- stan_glm(
+data_model <- stan_glm(
   formula = current_price ~ old_price + vendor + month,
-  data = beef_data,
+  data = data_parquet,
   family = gaussian(),
   prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
   prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
@@ -33,33 +32,33 @@ beef_model <- stan_glm(
 
 #### Save model ####
 saveRDS(
-  beef_model,
-  file = here("models", "beef_model.rds")
+  data_model,
+  file = here("models", "data_model.rds")
 )
 
 #### Visualization and Analysis ####
 
 # Chart 1: Distribution of Current Price
-ggplot(beef_data, aes(x = current_price)) +
+ggplot(data_parquet, aes(x = current_price)) +
   geom_histogram(binwidth = 1, fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
   geom_density(aes(y = ..count..), color = "red", size = 1) +
   labs(title = "Distribution of Current Prices", x = "Current Price (in $)", y = "Frequency") +
   theme_minimal()
 
 # Chart 2: Price Difference by Vendor
-beef_data <- beef_data %>%
+data_parquet <- data_parquet %>%
   mutate(price_difference = old_price - current_price)
 
-ggplot(beef_data, aes(x = vendor, y = price_difference)) +
+ggplot(data_parquet, aes(x = vendor, y = price_difference)) +
   geom_boxplot(fill = "lightgrey", color = "darkgrey", alpha = 0.7) +
[…]

### 2024-12-04 scripts/06-model_data.R
Update 06-model_data.R

@@ -1,28 +1,33 @@
-```r
 #### Preamble ####
-# Purpose: To build and save a Bayesian regression model that predicts the flying time of planes based on length and width using the `rstanarm` package with specified priors and a Gaussian family.
+# Purpose: Build and save a Bayesian regression model predicting the flying time of planes based on length and width using the `rstanarm` package with specified priors and a Gaussian family.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: 
-  # - The `rstanarm` and `tidyverse` packages must be installed.
-  # - The cleaned data file must be available in the specified path.
-  # - An organized project structure must be in place with a "models" directory.
-# Any other information needed? Make sure all dependencies are installed and paths are correct.
+#   - The `rstanarm` and `tidyverse` packages must be installed.
+#   - The cleaned data file must be available in the specified path.
+#   - An organized project structure must be in place with a "models" directory.
 
 #### Workspace setup ####
 # Load required libraries
 library(tidyverse)
 library(rstanarm)
+library(here)
 
 #### Read data ####
+# Define the path to the analysis dataset
+data_path <- here("data", "analysis_data", "analysis_data.csv")
+
 # Load the analysis dataset
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+if (!file.exists(data_path)) {
+  stop("The file 'analysis_data.csv' does not exist. Please ensure the data file is available.")
+}
+analysis_data <- read_csv(data_path)
 
 #### Model data ####
 # Build the Bayesian regression model
-first_model <-
+plane_model <-
   stan_glm(
     formula = flying_time ~ length + width,
     data = analysis_data,
@@ -34,9 +39,12 @@ first_model <-
   )
 
 #### Save model ####
+# Define the path to save the model
+model_path <- here("models", "plane_model.rds")
+
 # Save the model to the models directory
+dir.create(dirname(model_path), recursive = TRUE, showWarnings = FALSE)
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  plane_model,
+  file = model_path
 )
-``` 

### 2024-12-04 scripts/07-replications.R
Update 07-replications.R

@@ -1,4 +1,3 @@
-```r
 #### Preamble ####
 # Purpose: To replicate and visualize graphs from the analysis of plane data using the cleaned dataset.
 # Author: Duanyi Su
@@ -17,4 +16,3 @@ library(tidyverse)
 #### Load data ####
 # Load the cleaned data for analysis and visualization
 cleaned_data <- read_csv("data/analysis_data/analysis_data.csv")
-```

### 2024-12-04 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,6 +1,6 @@
 #### Preamble ####
 # Purpose: Simulate and save a dataset of Canadian grocery store products, including vendor,
-#          product details, and pricing information for analysis and testing.
+#          product details, and pricing information for analysis.
 # Author: Duanyi Su
 # Date: 3 December 2024
 # Contact: duanyi.su@mail.utoronto.ca
